Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens

### 2014-01-28 limits/ms.tex
New plots showing examples of no symmetry and changing pixrad.
Fixed up notation for inertia tensor.

The lens equation:
\begin{equation}
\begin{aligned}
\vec\beta &= \vec\theta - \frac{D_{LS}}{D_S}\vec\alpha(\vec\theta) \\
\vec\alpha(\vec\theta) &= \frac{4G}{c^2D_L} \int \Sigma(\vec\theta')
                          \frac{(\vec\theta - \vec\theta')}
                          {\ |\vec\theta - \vec\theta'|^2} \, d^2\vec\theta'
\end{aligned}
\label{eqn:lens_equation}
\end{equation}
maps an observed image position $\vec\theta$ to a source position
$\vec\beta$.  The lens can be thought of as a projected surface
density $\Sigma$ which diverts the path of a photon instantaneously
through the bending angle $\vec\alpha$.  The $D$ factors, as in the previous section, are angular
diameter distances, which depend on the particular choice of cosmological
parameters $\Omega$:
\begin{equation}
D_{LS} = \frac c{H_0} \frac{1+z_S}{1+z_L} \int_{z_L}^{z_S}
                      \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
%
and $D_L \equiv D_{0,L}$, $D_S \equiv D_{0,S}$.
One way to understand the lens equation is via Fermat's principle. We
can think of light as travelling only along extremum paths where
lensed images occur. Such paths occur at
the extrema of the photon {\it arrival time\/} $t(\vec\theta)$
that depends on the geometric path the photon takes and the general
relativistic gravitational time dilation due to a thin lens at
redshift $z_L$:
\begin{equation}
\begin{aligned}
\frac{ct(\vec\theta)}{(1+z_L)D_{L}}
&= {\textstyle\frac12} |\vec\theta - \vec\beta|^2
   \times \frac{D_{S}}{D_{LS}} \\
&- \frac{4GD_L}{c^2}
   \int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| \, d^2\vec\theta'
\label{full arrival time}
\end{aligned}
\end{equation}
We can simplify the above equation by introducing a dimensionless time $\tau$ and density $\kappa$:

@@

We can also more directly probe the recovery of the shape of the mass
distribution by defining a global measure of lens shape. We define this to be the ratio of the eigenvalues of the 2D inertia tensor:

\begin{equation} 
I_{ij} = \sum_\theta M(\vec\theta) \theta_i\theta_j \,\,\, ; \,\,\, s \equiv \lambda_1/\lambda_2
\label{eqn:inertia}
\end{equation}
where $\lambda_1$ is always taken to be the largest value. As with $f_R$ and $f_\theta$, we only consider mass up to one radial position passed the outermost image and compute the fractional error as:
%
\begin{equation} \label{ferror shape}
  f_\mathrm{shape} = \left|s - \tilde s\right|  / \tilde s
\end{equation}
%
where $\tilde s$ is the shape of the mock galaxy. The distribution of $f_\mathrm{shape}$
for each mock galaxy and each test case is shown in \figref{shape results}. Interestingly, for this global shape parameter recovery it appears more important to have time delay data and/or a central image (T/M) than to have a quad or multiple sources with wide redshift separation. In all cases, the stellar mass little-aids the recovery, reflecting the fact that $s$ is heavily weighted towards the shape at the {\it edge} of the mass map, rather than at the centre where the stars may dominate the potential (see equation \ref{eqn:inertia}).

@@

>>>>

The lens equation:
%
\begin{equation}
\begin{aligned}
    \vec\beta &= \vec\theta - \frac{D_{LS}}{D_S}\vec\alpha(\vec\theta) \\
\vec\alpha(\vec\theta) &= \frac{4G}{c^2D_L} \int \Sigma(\vec\theta')
                          \frac{(\vec\theta - \vec\theta')}
                          {\ |\vec\theta - \vec\theta'|^2} \, d^2\vec\theta'
\end{aligned}
\label{eqn:lens_equation}
\end{equation}
%
maps an observed image position $\vec\theta$ to a source position
$\vec\beta$.  The lens can be thought of as a projected surface
density $\Sigma$ which diverts the path of a photon instantaneously
through the bending angle $\vec\alpha$.  The $D$ factors, as in the previous section, are angular
diameter distances, which depend on the particular choice of cosmological
parameters $\Omega$:
\begin{equation}
D_{LS} = \frac c{H_0} \frac{1+z_S}{1+z_L} \int_{z_L}^{z_S}
                      \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
%
and $D_L \equiv D_{0,L}$, $D_S \equiv D_{0,S}$.
One way to understand the lens equation is via Fermat's principle. We
can think of light as travelling only along extremum paths where
lensed images occur. Such paths occur at
the extrema of the photon {\it arrival time\/} $t(\vec\theta)$
that depends on the geometric path the photon takes and the general
relativistic gravitational time dilation due to a thin lens at
redshift $z_L$:
\begin{equation}
\begin{aligned}
\frac{ct(\vec\theta)}{(1+z_L)D_{L}}
&= {\textstyle\frac12} |\vec\theta - \vec\beta|^2
   \times \frac{D_{S}}{D_{LS}} \\
&- \frac{4GD_L}{c^2}
   \int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| \, d^2\vec\theta'
\label{full arrival time}
\end{aligned}
\end{equation}
We can simplify the above equation by introducing a dimensionless time $\tau$ and density $\kappa$:

@@

We can also more directly probe the recovery of the shape of the mass
distribution by defining a global measure of lens shape
%
\begin{equation} 
    s \equiv \lambda_1/\lambda_2
\end{equation} 
%
where $\lambda_1$ and $\lambda_2$ are the eigenvalues of the 2D inertia tensor
%
\begin{equation} 
    I_{ij} = (-1)^{1-\delta_{ij}}\sum_\theta M(\vec\theta) \theta_i\theta_j 
\label{eqn:inertia}
\end{equation}
%
where $\delta_{ij}$ is the Kronecker delta. We always take $\lambda_1$ to be the largest value. As with $f_R$ and $f_\theta$, we only consider mass up to one radial position passed the outermost image and compute the fractional error as:
%
\begin{equation} \label{ferror shape}
  f_\mathrm{shape} = \left|s - \tilde s\right|  / \tilde s
\end{equation}
%
where $\tilde s$ is the shape of the mock galaxy. The distribution of
$f_\mathrm{shape}$ for each mock galaxy and each test case is shown in
\figref{shape results}.  Interestingly, for this global shape parameter
recovery it appears more important to have time delay data and/or a central
image (T,M) than to have a quad or multiple sources with wide redshift
separation. In all cases, the stellar mass little-aids the recovery, reflecting
the fact that $s$ is heavily weighted towards the shape at the {\it edge} of
the mass map, rather than at the centre where the stars may dominate the
potential (see equation \ref{eqn:inertia}).

@@

\section{No symmetric prior}
\begin{figure*}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-nosymm_Tms-kappa.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-nosymm_Tms-a.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-nosymm_Tms-b.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-nosymm_Tms-c.pdf} \\
\includegraphics[width=0.24\textwidth]{BCExtendedDoubleR1-nosymm_tms-kappa.pdf}
\includegraphics[width=0.24\textwidth]{BCExtendedDoubleR1-nosymm_tms-a.pdf}
\includegraphics[width=0.24\textwidth]{BCExtendedDoubleR1-nosymm_tms-b.pdf}
\includegraphics[width=0.24\textwidth]{BCExtendedDoubleR1-nosymm_tms-c.pdf}
\caption{No symmetry prior.}
\label{}
\end{figure*}

\section{Pixel resolution convergence test}
\begin{figure*}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr10_TmS-a.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr10_TmS-b.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr10_TmS-c.pdf} \\
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr7_TmS-a.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr7_TmS-b.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr7_TmS-c.pdf} \\
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr6_TmS-a.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr6_TmS-b.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr6_TmS-c.pdf} \\
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr5_TmS-a.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr5_TmS-b.pdf}
\includegraphics[width=0.24\textwidth]{BCQuadR1a-pr5_TmS-c.pdf} \\
\caption{Changing pixrad. top to bottom: pixrad=10,7,6,5. We use 8 normally.}
\label{}
\end{figure*}
